Report on a Safety Checker for Petri nets

### 2013-11-21 paper/workspace/main.tex
Updated paper
Added method Invariant Minimization to paper draft

\usepackage{amsmath}
\usepackage[usenames,dvipsnames]{color}
\usepackage{multirow}
\usepackage{listings}
\usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}

\usepackage{tikz}
\usetikzlibrary{arrows,shapes,shapes.multipart,shapes.geometric,
	snakes,automata,backgrounds,petri,calc}
\usetikzlibrary{positioning}

@@

\begin{center}
  \begin{tikzpicture}
    \node[state] (begin) {BEGIN};
    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
    \node[print, right=of satc] (noinv) {No Invariant};
    \node[action, below=of satc] (inv) {$A:=\text{Model}(C')$\\
      $I := \text{Inv}(N, P, A)$};
    \node[print, below=of inv] (printinv) {Invariant: $I$};
    \node[state, right=of noinv] (end1) {END};
    \node[state, below=of printinv] (end2) {END};

@@

  Input:
    (S, T, E, I) : Petri net
    P            : Property
  Output:
    C'           : Dual state constraints

Input:
 N = (S, T, E, I)
   = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
 P = ( cs1 + cs2 <= 1 )

@@

  Input:
    (S, T, E, I) : Petri net
    A            : Satisfying assignment for C'

@@

\newpage
\section{Method Safety by Refinement}

@@

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | r | p{7cm} | r | r | r | r | } %
    \hline
    \multicolumn{2}{|l|}{Configuration} & pos & neg & don't know & TO \\
    \hline
    C1 & State constraint over real numbers & 23 &  0 & 54 &  0 \\
    C2 & State constraint over integers        & 23 &  0 & 54 &  0 \\
    \hline
    C3 & C2 + TrapConditions                   & 23 &  0 & 54 &  0 \\
    C4 & C3 + SubnetTrapConditions         & 23 &  0 & 50 &  4 \\
    C5 & C4 + EmptyTrapConditions          & 23 &  0 & 46 &  8 \\
    \hline
    C6 & C1 + state space exploration \& TO 1\,min & 26 & 34 & 0 & 17 \\
    C7 & C1 + state space exploration \& TO 10\,min & 26 & 39 & 0 & 12 \\
    C8 & C1 + state space exploration \& TO 8\,h   & 26 & 45 & 0 &  6 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \bfc}
\label{bfc-experiments}
\end{table}

@@

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | r | p{6cm} | r | r | r | r | }
    \hline
    \multicolumn{2}{|l|}{Configuration} & pos & neg & don't know & TO \\
    \hline
    C1 & State constraint with real numbers & 14 &  0 & 13 &  0 \\
    C2 & State constraint with integers     & 14 &  0 & 13 &  0 \\
    \hline
    C3 & C2 + TrapConditions               & 20 &  0 &  7 &  0 \\
    C4 & C3 + SubnetTrapConditions         & 21 &  0 &  6 &  0 \\
    C5 & C4 + EmptyTrapConditions          & 21 &  0 &  6 &  0 \\
    \hline
    C9 & C4 + state space exploration \& TO 4\,h & 21 &  2 & 0 &  4 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \mist}
\label{mist-experiments}
\end{table}

@@

Invariant $\text{Inv}(N, P, A) = (cs1 + cs2 + m \le 1)$

\paragraph{Technique}
\begin{align*}
  \mathcal C'(N, P) =& ( ( ( Y_1 \cdot A + Y_2) \cdot C \le 0) \land \\
& ((Y_1 \cdot A + Y_2) \cdot M_0 < Y_1 \cdot b) \land \\
& (Y_1 \ge 0) \land (Y_2 \ge 0)) \\
  \text{Inv}(N, P, A) =& ( ( Y_1 \cdot A + Y_2) \cdot M \le ( Y_1 \cdot A + Y_2) \cdot M_0)
\end{align*}
where $M_0$ is the initial marking, $C$ is the incidence matrix of the net and
$A\cdot M \ge b$ is equivalent to $\neg P$.

\paragraph{Method Invariant}
\mbox{ } \\
\begin{tikzpicture}
  \node[state] (begin) {BEGIN};
  \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
  \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
  \node[print, right=of satc] (noinv) {No Invariant};
  \node[action, below=of satc] (inv) {$A_\theta:=\text{Model}(C')$\\
                                  	$I := \text{Inv}(N, P, A)$};
  \node[print, below=of inv] (printinv) {Invariant: $I$};
  \node[state, right=of noinv] (end1) {END};
  \node[state, below=of printinv] (end2) {END};

>>>>

\usepackage{amsmath}
\usepackage[usenames,dvipsnames]{color}
\usepackage{listings}
\usepackage[a4paper, top=2cm, bottom=3cm, left=2cm, right=2cm]{geometry}

\usepackage{tikz}
\usetikzlibrary{arrows,shapes,shapes.multipart,shapes.geometric,
	snakes,automata,backgrounds,petri,calc,positioning}

@@

\begin{center}
  \begin{tikzpicture}
    \node[state] (begin) {BEGIN};
    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
    \node[print, right=of satc] (noinv) {No Invariant};
    \node[action, below=of satc] (inv) {$A:=\text{Model}(C')$\\
      $I := \text{Inv}(N, A)$};
    \node[print, below=of inv] (printinv) {Invariant: $I$};
    \node[state, right=of noinv] (end1) {END};
    \node[state, below=of printinv] (end2) {END};

@@

  Input:
    (S, T, E, M0) : Petri net
    P             : Property
  Output:
    C'            : Dual state constraints

Input:
 N = (S, T, E, M0)
   = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
 P = ( cs1 + cs2 <= 1 )

@@

  Input:
    (S, T, E, M0) : Petri net
    A             : Satisfying assignment for C'

@@

\newpage
\section{Method Invariant Minimization}

The method Invariant Minimization constructs an invariant \verb=I= that uses a minimal number places for given Petri net \verb=N= and \verb=P= when \verb=N= never violates \verb=P=.
We present the method Invariant Minimization by example on a simplified mutual exclusion algorithm by Delzanno2001.

  Subprocedure \text{MinConstraints} constructs minimization constraints C_M corresponding to N and A
\end{verbatim}

\begin{center}
  \begin{tikzpicture}
    \node[state] (begin) {BEGIN};
    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
    \node[print, right=of satc] (noinv) {No Invariant};
    \node[action, below=of satc] (model1) {$A:=\text{Model}(C')$};
    \node[action, below=of model1] (cm) {$C_M:=\text{MinConstraints}(N, A)$};
    \node[decision, below=of cm] (satcm) {$\text{SAT}(C' \cup C_M)$};
    \node[action, right=of satcm] (model2) {$A:=\text{Model}(C' \cup C_M)$};
    \node[action, below=of satcm] (inv) {$I := \text{Inv}(N, A)$};
    \node[print, below=of inv] (printinv) {Invariant: $I$};
    \node[state, right=of noinv] (end1) {END};
    \node[state, below=of printinv] (end2) {END};

    \draw (begin) edge (c);
    \draw (c) edge (satc);
    \draw (satc) edge node[above]{NO} (noinv);
    \draw (noinv) edge (end1);
    \draw (satc) edge node[right]{YES} (model1);
    \draw (model1) edge coordinate[pos=.5] (edgein) (cm);
    \draw (cm) edge (satcm);
    \draw (satcm) edge node[above]{YES} (model2);
    \draw (model2) |- (edgein);
    \draw (satcm) edge node[right]{NO} (inv);
    \draw (inv) edge (printinv);
    \draw (printinv) edge (end2);
  \end{tikzpicture}
\end{center}

\begin{verbatim}
* Property of minimization constraints C_M generated from A: 
    If A' satisfies C' u C_M, then Inv(N, A') uses less places than Inv(N, A)
\end{verbatim}

\newpage
\begin{verbatim}
* Example:
  - Dual state constraints C' and satisfying assignment A for C' as in example for method Invariant

  - Minimization constraints:

    nc  ~= 0 => m_nc = 1
    |           --------
    |              |
    |           place appears in invariant
    |
    place coefficient in the invariant

    nc   = 0 => m_nc  = 0
    |           ---------
    |              |
    |           place does not appear in invariant
    |
    place coefficient in the invariant

    m   ~= 0 => m_m   = 1
    m    = 0 => m_m   = 0
    cs1 ~= 0 => m_cs1 = 1
    cs1  = 0 => m_cs1 = 0
    cs2 ~= 0 => m_cs2 = 1
    cs2  = 0 => m_cs2 = 0

    m_nc + m_m + m_cs1 + m_cs2 < 3
    --------------------------   |
                |                |
                |    number of places appearing in current invariant for A
                |    = #{s | A(s) ~= 0} = #{m, cs1, cs2}
                |
                |
  number of places appearing in new invariant
\end{verbatim}

\begin{verbatim}
* Subprocedure \text{MinConstraints}:

  Input:
    (S, T, E, M0) : Petri net
    A             : Assignment
  Output:
    C_M           : Minimization Constraints

Input:
 N   = (S, T, E, M0)
     = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
 A   = ( nc=0, m=1, cs1=1, cs2=1, target1=1 )

\newpage

@@

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | r | p{7cm} | r | r | r | r | } %
    \hline
    \multicolumn{2}{|l|}{Configuration} & pos & neg & don't know & TO \\
    \hline
    C1 & State constraint over real numbers & 23 &  0 & 54 &  0 \\
    C2 & State constraint over integers        & 23 &  0 & 54 &  0 \\
    \hline
    C3 & C2 + TrapConditions                   & 23 &  0 & 54 &  0 \\
    C4 & C3 + SubnetTrapConditions         & 23 &  0 & 50 &  4 \\
    C5 & C4 + EmptyTrapConditions          & 23 &  0 & 46 &  8 \\
    \hline
    C6 & C1 + state space exploration \& TO 1\,min & 26 & 34 & 0 & 17 \\
    C7 & C1 + state space exploration \& TO 10\,min & 26 & 39 & 0 & 12 \\
    C8 & C1 + state space exploration \& TO 8\,h   & 26 & 45 & 0 &  6 \\
    \hline
    C10 & Method Invariant                  & 23 &  0 & 54 &  0 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \bfc}
\label{bfc-experiments}
\end{table}

@@

\begin{table}[h]
\begin{center}
  \begin{tabular}{ | r | p{6cm} | r | r | r | r | }
    \hline
    \multicolumn{2}{|l|}{Configuration} & pos & neg & don't know & TO \\
    \hline
    C1 & State constraint with real numbers & 14 &  0 & 13 &  0 \\
    C2 & State constraint with integers     & 14 &  0 & 13 &  0 \\
    \hline
    C3 & C2 + TrapConditions               & 20 &  0 &  7 &  0 \\
    C4 & C3 + SubnetTrapConditions         & 21 &  0 &  6 &  0 \\
    C5 & C4 + EmptyTrapConditions          & 21 &  0 &  6 &  0 \\
    \hline
    C9 & C4 + state space exploration \& TO 4\,h & 21 &  2 & 0 &  4 \\
    \hline
    C10 & Method Invariant                  & 14 &  0 & 13 &  0 \\
    \hline
  \end{tabular}
\end{center}
\caption{Experimental results for the benchmarks of \mist}
\label{mist-experiments}
\end{table}

@@

Invariant $\text{Inv}(N, A) = (cs1 + cs2 + m \le 1)$

\paragraph{Technique}
\begin{align*}
  \mathcal C'(N, P) =& ( ( ( Y_1 \cdot A + Y_2) \cdot C \le 0) \land \\
& ((Y_1 \cdot A + Y_2) \cdot M_0 < Y_1 \cdot b) \land \\
& (Y_1 \ge 0) \land (Y_2 \ge 0)) \\
  \text{Inv}(N, A) =& ( ( Y_1 \cdot A + Y_2) \cdot M \le ( Y_1 \cdot A + Y_2) \cdot M_0)
\end{align*}
where $M_0$ is the initial marking, $C$ is the incidence matrix of the net and
$A\cdot M \ge b$ is equivalent to $\neg P$.

\paragraph{Method Invariant}
\mbox{ } \\
\begin{tikzpicture}
  \node[state] (begin) {BEGIN};
  \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
  \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
  \node[print, right=of satc] (noinv) {No Invariant};
  \node[action, below=of satc] (inv) {$A_\theta:=\text{Model}(C')$\\
                                  	$I := \text{Inv}(N, A)$};
  \node[print, below=of inv] (printinv) {Invariant: $I$};
  \node[state, right=of noinv] (end1) {END};
  \node[state, below=of printinv] (end2) {END};
